\documentclass[a4paper,11pt]{article}
\usepackage{graphicx,listings,a4wide,dsfont}
%\usepackage[firstpage]{draftwatermark}
%\SetWatermarkLightness{0.5}
%\SetWatermarkScale{4}
\setcounter{tocdepth}{2}

\newcommand{\question}[2]{\medskip\par\noindent\textbf{#1}\\\hangindent=0.5cm#2}

\title{Report on OGO 2.2 \\ Software specification\\ Review formal specification group 3}
\author{
        Tim van Dalen, Tony Nan, Ferry Timmers, \\ Lasse Blaauwbroek, Femke Jansen, \\Jeroen Peters and Sander Breukink\\ OGO 2.2 group 6 \\
                Department of Computer Science\\
        Technical University Eindhoven\\
}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This document contains the review of the formal specification of group 3. First some general remarks about the overall structure of the formal specification are provided in section 2. \\
In section 3, our opinion and remarks about the Z-specification are provided; the remarks are organized in a similar way to the structure of the formal specification. In sections 4 and 5, the state diagram and the MSC are evaluated. In each of these sections, the remarks are divided into two categories: general remarks and individual remarks. The general remarks apply to several parts of the specification; the individual remarks apply to individual methods or parts of the specification. Our remarks are provided as lists, in order to keep them organized and readable. \\
In section 6, our judgement and grading for the formal specification is given. The final grade consists of four partial grades of the following categories: consistency, correspondence to the informal specification, completeness and explanation \& coherence. For each of these partial grades, we list the issues that had a major impact on the grade.
\end{abstract}

\newpage
	
	\tableofcontents
	\newpage

    \section{Introduction}
    	This document contains the review for the formal specification described in "Formele specificatie van Groep 6 OGO 2.2" by group 3. 
	It describes the differences between the informal specification given in "Report on OGO 2.2 Software specification Assignment 2a".
	Any incosistencies between the specifications or within the formal specification itself will be listed, and any comments on the MSC, state chart or Z-specification will be given.
	The formal specification will be graded on consistency, correspondence to the informal description, completeness and explanation and coherence.

	\section{General remarks}
    \begin{itemize}
        \item The order of sections is not logical: the Z specification should be after the MSC's and the Statecharts.
        \item The abstract is missing.
        \item There is no table of contents.
        \item The introduction is on the short side and is actually more of an abstract (yet if it were the state diagrams and MSC's are not mentioned).
        \item There are no use case scenario's (and also no matching state diagrams and MSC's).
        \item The design decisions that have been made are not documented. For example, in our informal specification, it is not specified that player 1 always makes a turn before player 2; this decision should be documented and motivation.
        \item A class diagram would be appreciated.
        \item It is not specified how a game ends and how the individual components of the game (Controller, Board, etc.) can conclude this and react to this.
    \end{itemize}
	
	\section{Remarks about the Z-specification}
    In this section, our individual comments are organized in their respective order in the formal specification.
    \subsection{Overall remarks Z-specification}
    \begin{itemize}
        \item The description of your coordinate-system is very nice and understandable. However, the two picture are not consistent; they both describe a different coordinate-system. The second coordinate-system is used in the Z-specification, but the first coordinate-system is never used in the document.
        \item Dividing the "MoveRequest" in Controller in several parts was a wise decision; this way the reader is introduced step-by-step to the whole concept.
        \item Boolean is a pre-defined set in Z. The correct notation for a set of booleans is $\mathds{B}$.
        \item In sections 2 to  3.5, input or output variables are \emph{pre}fixed with a ? or !. In the remaining sections of the Z-specification, the input or output variables are \emph{post}fixed with a ? or !. This is inconsistent and partly incorrect; input or output variables should be \emph{post}fixed with a ? or !.
        \item Some variables are unbounded or even undefined. For example, in the schema DoKill, a variable $s$ is used, but it is not clear where this variable comes from and what its type is. The same holds for the $s?$ variable in DoMove.
        \item The MoveRequest in player checks whether a move is possible, but that is not the job of the player; the board should check whether a move request is valid.
        \item The notion of "safe tiles" is not reflected in the Z-specification. This is an important factor when a fox wants to kill a dolphin and vice versa.
        \item The notion of rounds is explained informally, but not reflected formally in the Z-specification. For example, the tides occur after completion of a round.
        \item A Number of Moves schema is specified, but not used in MoveRequest or DoMove. A move can only be requested (and executed) if a player has any moves left.
        \item The board has a size that depends on the number of pieces. This is not reflected in the formal specification.
        \item It is not always clear where the input, output and dummy variables are used for in the Z-schemas. Additional explanation would be appreciated.
	   \item Some Z functions are being used before they are defined. For example, the function Neighbours is used in Board, but it is defined one page later. In the informal explanation, there is no reference to this later defined function.
        \item Sometimes in function, the output comes first and sometimes the input. This should be consistent.
        \item In section 2, Pieces is defined as a schema with an attribute 'type'. This attribute is also used in the informal explanation in the sections 2 to 3.3.3. However, in section 3.4 till 4.2, this attribute is referred to as 'species'; this is inconsistent and even undefined. Later on, in section 4.2 till the end of the document, the attribute is again referred to as 'type'.
   \end{itemize}

    \subsection{Pieces and players}
    \begin{itemize}
        \item The document only specifies that each player has pieces of the same type; one player should have pieces of all type 'fox' and one player should have pieces of all type 'dolphin'.
        \item All pieces have type animal; it would better to simply use animals in stead of pieces, in our opinion.
        \item A piece can move beyond the borders of the board.
        \item An animal can kill an animal of the same species.
        \item 'Pieces' is not defined yet; it is used in the DoKill function among others. However 'Piece' is defined.
        \item The MoveRequest function does not use the neighbors function.
        \item The DoMove function and MakeMove (defined at Board) look ambiguous.
    \end{itemize}

    \subsection{Board}
    \begin{itemize}
        \item The default water height invariant is a bit confusing. First, it is stated that the $x$ value of a nEvent must be between $-20$ and $20$, however the invariant says that the $x$ and $y$ variables of nEvent are both natural numbers.
        \item The natural events in the scheme NaturalEvent should be modeled with a random factor. In the formal specification, it it specified that whenever a natural events ends, a new natural event will immediately begin. This is not specified in the informal specification.
        \item In the informal explanation of Tide, it is stated that the change of the tide is uniformly distributed for any number between -5 and 5. This is however not specified formally in the Z specification.
        \item In the informal explanation of Flood, it is stated that the Flood function is called twice; however, this is not clearly reflected in the Z-schema.
        \item In OccupiedBySameAnimal, the Z-schema is correct, but the informal explanation is not. It states the following: if the destination tile of a move request is occupied by an animal of the same type, then isOccupied is true and the move to this tile is possible. In this case the move is, of course, not possible.
        \item In MakeMove, a conjunction is missing between the parts of the fox and the dolphin.
        \item In ShortcutPossible, both the dolphins and foxes use the bridges now as a shortcut. Dolphins should, of course, use the caves.
        \item In the informal explanation of getNrMoves, it is mentioned that the operation calculates the number of moves "automagically", yet it does not state what this means.
        \item A bridge or tunnel can begin and end on the same tile or begin on one tile and end on an adjacent tile.
        \item The terms 'days' and 'turns' are used as if they mean the same thing, whereas 'days' should indicate how many turns there are in one day.
        \item In the inRange function, p.species is used, but only p.type is defined.
        \item In the MakeMove and ShortcutPossible functions, a piece.type is given where a piece is required.
    \end{itemize}

    \subsection{Viewer and Controller}
    \begin{itemize}
        \item In Init, two foxes and two dolphins should not be placed on the same tile. This is captured in MoveRequest, but it must also be captured in the initial configuration.
        \item The initial number of foxes and dolphins can be predefined by the players. It is not clear whether this is the $n?$ variable in init; if it is, explain this in the informal description.
        \item In Init, all foxes should be placed on the land and all dolphins in the water.
        \item The Init function allows players to begin with $0$ pieces, which is of course an exceptional case that should be taken care of.
    \end{itemize}

	\section{Remarks about the state diagram}
    \begin{itemize}
        \item The order of the informal explanation is inconsistent with the order of explanation in the Z-specification. In the Z-specification, the Z-schema is provided first and thereafter the explanation; in the state diagram and MSC, the explanation is provided first and thereafter the diagrams. This is inconsistent and also confusing.
        \item The state diagrams of the individual components are missing. The individual state diagrams reflect the communication within an entity (internal actions) and to the entities it "knows about"; the system state diagram is used to show how the individual entities communicate with each other.
        \item In the state diagram, it is reflected that there always is an initial flood, after the board and all its components have been initialized. There always is a flood after a round has been completed; not necessarily before the first round.
        \item A reference from a state diagram to a message sequence chart should be avoided. For every message sequence chart, there should be a state diagram which correspondents to it.
        \item All state diagrams should have a terminating state, because the game can end. Also, some classes like the player-classes should have an initializing state to indicate that they are being initialized.
        \item A lot of transition labels are incorrect. For instance, the label "MSC for P1 / End P1" of Player 1 is the part "/ End P1" not needed, since the transition "/ P2" can only be executed if state "Turn Completed" is active.
        \item There is a transition in View that does not have a label. This could either mean that it is executed immediately after the state "Updated" becomes active, or that it will be randomly executed. Explain what this transition means. It could also mean that this transition is useless.
        \item Many triggers in transitions are not in the form of a method-call, like "NEvent" in Main. An explanation what this label means is missing; for example, that it is a notification that is send from one class to another
        \item The triggers of the transition in View are incorrect. According to the current state diagram, the update of the view and the update of the flood are executed concurrently, however it seems more logical that they execute sequentially.
        \item A part of the following sentence appears to be missing: "Na een beurt van een Player wordt ook het View weer up-to-date gemaakt, zoals."
    \end{itemize}

    \section{Remarks about Message Sequence Chart}
    \begin{itemize}
        \item The diagram is too small; it is very difficult to read it properly when printed.
        \item In the MSC is shown that a round begins with a call from the controller to the board, to request the number of moves of a player. This is in conflict with our informal specification. A player makes a move request first; then the controller forwards this request to the board. Since the board knows the number of moves for each player it can internally determine whether a move request is valid based upon these number of moves.
        \item The "request move"-message from controller to player should go the other way around. As explained above, the player requests a move; the controller does not ask the player whether it wants to move.
    \end{itemize}


    \section{Judgement and grading}
    \subsection{Consistency}
    There are a few inconsistencies in the report: for example, the attribute 'type' in Player is sometimes referred to as 'species' and sometimes as 'type'. Due to the fact that the state charts and MSC's were not complete enough, these could not be checked for consistency. Also, the two coordinate-systems in section 3.2.1 are not consistent with each other. Based upon these issues, the grade for consistency a 1.5 out of 2.5.\\

    \subsection{Correspondence to the informal specification}
     Some things stated in the informal description are not mentioned in the formal specification. For example, one player should control all the foxes, while the other player should control all the dolphins. Also, an important factor that is not reflected in the formal specification is the notion of a "safe tile". The rest of the formal specification does correspond to the informal description; the grade for correspondence is therefore a 1.5 out of 2.5. \\

    \subsection{Completeness}
    Since there are no abstract and introduction and the MSC and Statechart are not sufficient the grade for this part is 0.5.\\

    \subsection{Explanation and coherence}
    The explanation matches with the functions pretty well, therefor the grade of this part is a 2.3.\\

    \subsection{Conclusion}
    The total grade is equal to 1.5 + 1.5 + 0.5 + 2.3 = 5.8.\\

\end{document}
