This section contains the message sequence charts, both of high and class level.
\subsection{High Level Message Sequence Chart}
	
	\begin{minipage}{\linewidth}
		The following graph represents our high level message sequence chart and shows how a normal program flow is modeled by MSCs.
		
		\input{hmsc}	
	\end{minipage}
	\newpage
\subsection{Message Sequence Charts}
	This section contains the message sequence charts for our use cases. Below every MSC is the location of the MSC in the High Level Message Sequence Chart, which shows the possible routes that can be taken from the corresponding MSC. For example, below the MSC 'Initialize', both 'Initialize' and 'Move Request' are shown, since 'Move Request' is the next step after 'Initialize' has been completed. If several options for next moves exist, they are given in the same way as in the HMSC below; in the corresponding separate MSC is explained when and why a certain route is taken. Note that the below MSCs are identical to the use cases described in section 1.

	The world entity is not a real part of our MSCs, but rather a link to the outside world. When an internal action is postfixed with '()' it's a function call to a private function of the entity. Otherwise, it is an action within the called function.

	The following MSC syntax is used in this section:

	\input{MSC-files/uitleg.tex}

	\subsubsection{Initialize}
	\begin{minipage}{\linewidth}
		When the game starts, the board is initialized. When this is done, the board sends a preInitialize to the controller. When the controller is done with that, the board will initialize all robots (choosing the appropriate rule for them), which will reply with an 'OK' when done. When all robots have been initialized, the board sends a postInitialize to the controller. All entities are now initialized. \\
		Separate pre- and post-initialize method are used because the controller can not fully initialize without the robots and the robots, in turn, must be initialized with the controller. So, the pre-initialize is used to create the controller and the post-initialize is used to store all the robots that have been initialized by the board.

		\input{MSC-files/initialize.tex}
    \end{minipage}

    \subsubsection{Initialize viewer}
	\begin{minipage}{\linewidth}
		The viewer is initialized by the world. Note that the world provides a parameter 'controller' to the viewer, so the viewer knows how to address the controller. Analogous to the HMSC, this happens after the components of the game (board, controller, etcetera) have been initialized. After the call has been received, the viewer will initialize itself and will attach itself to the controller, via addViewer. Once the viewer has been succesfully attached, the controller returns itself to notify the viewer of this success.
	
		\input{MSC-files/viewer_initialize.tex}
	\end{minipage}

	\subsubsection{Robot move request}
	\begin{minipage}{\linewidth}
		A robot can make a move request with the controller, which will forward that to the board. The board will check the validity of the move itself with the robots rule and then check the validity of the move at the current state of the board. The internal actions 'get possible moves' and 'get possible rotations' are thus used by the board to check whether the move request is conform with the specified moves and rotations in the rule of the robot. Based upon the outcome of the move request, one of the five MSCs given in the HMSC below is executed. \\
		Note that the move requests describe a move request in terms of local coordinates, since a robot does not know its exact location on the board. Also, the move request contains the robot that requested the move, along with the requested rotation.

		\input{MSC-files/robotmove_request.tex}
    \end{minipage}

	\subsubsection{Return Normal tile}
	\begin{minipage}{\linewidth}
		If the move request is valid, and the robot is moved to a normal tile, the board will return \emph{SUCCESS} to the controller, which will forward this message to the robot that was moved. Internally, the board calculates the new location of the robot and saves the location if the move request is valid.

		\input{MSC-files/robotmove_to_normal.tex}
	\end{minipage}

	\subsubsection{Return Hint tile}
	\begin{minipage}{\linewidth}
		If the move request is valid, and the robot is moved to a hint tile, the board will first respond with \emph{SUCCESS}, to indicate that the move request has been granted. Hereafter, the board generates a hint then notifies the controller that the robot that moved should receive a hint; the controller will forward the hint to the robot. Note that the notifyHint from the board to the controller also contains a robot, since the controller must be able to identify the robot that should receive the hint.

		\input{MSC-files/robotmove_to_hint.tex}
	\end{minipage}

	\subsubsection{Return Conveyor tile}
	\begin{minipage}{\linewidth}
		If the move request is valid, and the robot is moved to a conveyor tile, the board again returns \emph{SUCCESS} to the controller, which forwards this message to the robot that was moved. The board will then notify the controller that the robot was moved successfully and the controller will forward this to the robot. The board will then send a message to the controller that the robot was moved automatically (this was due to the conveyor belt) and the controller will forward this message to the robot, using notifyAutoMovement. As in the 'Return Hint tile' MSC, the notifyAutoMovement from the board to the controller contains the robot that should be notified.

		\input{MSC-files/robotmove_to_conveyor.tex}
	\end{minipage}

	\subsubsection{Return Home tile}
	\begin{minipage}{\linewidth}
	   If the move request is valid, and the robot is moved to its home tile, the board will notify the controller that the robot that moved wins the game, using the response \emph{WIN}.

		\input{MSC-files/robotmove_to_home.tex}
	\end{minipage}

	\subsubsection{Reject move}
	\begin{minipage}{\linewidth}
		If the move request is, for whatever reason, not valid, the board will notify the controller of this, using the response \emph{FAILED}. The controller will forward this to the robot that made the move request.

		\input{MSC-files/robotmove_reject.tex}
	\end{minipage}

	\advance\count17 by -6

    \subsubsection{Update viewer}
	\begin{minipage}{\linewidth}
        Whenever the board has changed, either because of a move request or a tiles exchange, the board prepares a snapshot for the viewer.	The board notifies the controller that it has a new snapshot for the viewer, using notifyViewer. The controller will then notify the viewer and send the board snapshot, using notifyStateChange; the viewer will internally update its view.

		\input{MSC-files/viewer_update.tex}
	\end{minipage}

	\subsubsection{Ordinary exchange}
	\begin{minipage}{\linewidth}
		The controller requests the board to do a tile exchange. The board will get two valid tiles, swap them and inform the controller that the tiles exchange was successful. Note that the 'getValidTiles'-function must maintain the "always reachable home tile"-invariant: each robot must always be able to reach its home tile. The board will calculate this internally and selects two valid tiles without robots in this case.

		\input{MSC-files/tiles_exchange.tex}
	\end{minipage}

	\subsubsection{Special exchange}
	\begin{minipage}{\linewidth}
		As in the ordinary exchange, the controller requests the board to do a tile exchange. The board will find two valid tiles, in this case with on at least one of them a robot or a conveyor belt, and swap them. The selected conveyor tiles and robots will be rotated. Once the tiles exchange is complete, the board will inform the controller. If one or two robots were part of the tiles exchange, the board will inform the controller of this; the controller will then inform the appropriate robots that they have been automoved, because of the exchange. \\
In the MSC below, one robot and one conveyor belt have been selected. Note that robot2 will not be notified, since it was not part of the tiles exchange. As in the MSC 'Ordinary exchange', the board must (internally) make sure that the "always reachable home tile"-invariant is maintained.

		\input{MSC-files/tiles_exchange_special.tex}
	\end{minipage}	

	\subsubsection{Notify robots}
	\begin{minipage}{\linewidth}
		The board first checks recursively for robots that are affected by the original move request. Then, the board sends a notification to the controller, which forwards this to the robot that was moved. This MSC is used when a robot is moved due to the move of another robot, for example when a robot that blocked the end of a conveyor belt moves away and the last robot on the belt should move to the end.
		Note that is it possible for a robot to end up on its home tile like this. Because of this, a BoardResponse is sent back, to indicate a winning (or of course normal) automove.

		\input{MSC-files/notify_robots.tex}
	\end{minipage}

	\subsubsection{End game}
	\begin{minipage}{\linewidth}
		If a robot has reached its home tile, the controller will send a terminate message to all losing robots, which will then terminate safely. The controller will then notify the viewer which robot won the game, so it can show the end game animation. When the animation is done, the viewer detaches itself from the controller, using removeViewer. The controller will then send a message to the board that everything it done and the board can reset and will then terminate. Note that the winning robot can terminate on its own, since its termination does not affect the other components of the game; the viewer can show an appropriate animation that reflects which robot has won. The board can then choose to either reset to start a new game or terminate; since this is a non-deterministic choice from the board, this is not modeled in the MSC below.

		\input{MSC-files/end_game.tex}
	\end{minipage}
